---
title: "Project Part 1"
author: "Rebeca De La Garza Evia"
date: "2022-11-06"
output: pdf_document
description: "Within this project, I use the Netflix Movies and TV Shows dataset to analyze the factors that result in longer runtimes measure either in minutes for movies or seasons for TV shows. Following this, I explore interesting questions that arise from the initial analysis."
---

Hello, before exploring the data, I first need to install a new packages and import the dataset itself. This is done in the following chunk. I use summary to ensure the data was imported properly and gain some basic information on the dataset. The head function lets me see how the data looks in the dataset.


```{r}
library(ggplot2)
library(tidyverse)
library(magrittr)
library(rlang)

netflix <- read.csv('Data//netflix_titles.csv')
summary(netflix)
head(netflix)
```

In this next chunk, I will begin my analysis by observing how runtime relates to the directors. I do this by first spliting the main dataset into two, one for movies and one for tv shows followed by determining which directors have the most works within each of these categories on Netflix and then observing their runtime data individually.

```{r}
movies <- subset(netflix, type =="Movie")
movies
shows <- subset(netflix, type =="TV Show")
shows
#makes two datasets


mcount <- count(movies, director)
mcount <- mcount[order(-mcount$n), ]
mcount
#creates a dataframe of the directors and how many netflix movies theyve directed

#scount <- count(shows, director)
#scount <- scount[order(-scount$n), ]
#scount
#creates a dataframe for show directors, unfortunetly show directors dont direct tons of shows like movie directors so this is useless 

topd <- c("Rajiv Chilaka", "Raúl Campos, Jan Suter",	"Suhas Kadav", "Marcus Raboy", "Jay Karas", "Cathy Garcia-Molina", "Jay Chapman", "Martin Scorsese", "Youssef Chahine")

#list of directors with the most movies

moviespop <- movies
topd
counter <-1

for (i in 1:6131){
  if (!(movies$director[i] %in% topd)){
    moviespop <- moviespop[-c(counter),]
  } else {
    counter <- counter+1
  }
}

#this makes a dataset of only the movies with the relevant directors

minutes <- c()

for (i in 1:131){
  addd <- unlist(str_split(moviespop$duration[i], " "))
  minutes <- append(minutes, addd[1])
}


moviespop$runtime <- minutes
moviespop
#here i created a new column with the runtimes but in integers

moviesdr<- data.frame(moviespop$director, moviespop$runtime)
#this new dataframe has just the directors and their respective movie runtimes
```

here i create a bar plot for the directors

```{r}
dirvrun <- ggplot(data = moviespop, aes(x = director, y = as.numeric(runtime), fill = director)) + geom_boxplot() + coord_flip() + theme(legend.position="none") + scale_fill_brewer(palette="BrBG")

dirvrun + ggtitle("Movie Director Boxplot") + xlab("Director") + ylab("Runtime")

#makes boxplot

moviesdr
tapply(as.numeric(moviesdr$moviespop.runtime), moviesdr$moviespop.director, summary)
#applys the summary for each director
```
for this next chunk i will begin to analyze runtime with regard to country of production

```{r}
mcountry_list <-c()
for (i in 1:6131){
  addd <- unlist(str_split(movies$country[i], ", "))
  mcountry_list <- append(mcountry_list, addd)
}
mcountry_list_df <- data.frame(mcountry_list)
#mcountry_list_df

mcountc <- count(mcountry_list_df, mcountry_list)
mcountc <- mcountc[order(-mcountc$n), ]
mcountc

#this creates a dataframe with each country individually and then creates a second dataframe that counts their occurances

scountry_list <-c()
for (i in 1:6131){
  addd <- unlist(str_split(shows$country[i], ", "))
  scountry_list <- append(scountry_list, addd)
}
scountry_list_df <- data.frame(scountry_list)
#scountry_list_df

scountc <- count(scountry_list_df, scountry_list)
scountc <- scountc[order(-scountc$n), ]
scountc

#this does the same thing but for the tv shows

mtopc <- c("United States", "India",	"United Kingdom", "Canada", "France", "Germany", "Spain", "Japan", "China")

#list of countries with the most movies

mcountrypop <- movies
mtopc
counter <-1

for (i in 1:6131){
  for (j in movies$country[i]){
    if (!(movies$country[i] %in% mtopc)){
      mcountrypop <- mcountrypop[-c(counter),]
  } else {
      counter <- counter+1
      break
  }
  }
}
mcountrypop

#this creates a dataframe with only the movies that were produced in one of the relevant countries

minutes <- c()

for (i in 1:3602){
  addd <- unlist(str_split(mcountrypop$duration[i], " "))
  minutes <- append(minutes, addd[1])
}


mcountrypop$runtime <- minutes
mcountrypop
#here i created a new column with the runtimes but in integers

moviesc<- data.frame(mcountrypop$country, mcountrypop$runtime)

#this new dataframe has just the countries and their respective movie runtimes

moviesc
```
this chunk will make the datasets for shows using the same methods

```{r}
stopc <- c("United States", "United Kingdom",	"Japan", "South Korea", "Canada", "France", "India", "Taiwan")

#list of countries with the most shows

scountrypop <- shows
stopc
counter <-1

for (i in 1:2676){
  for (j in shows$country[i]){
    if (!(shows$country[i] %in% stopc)){
      scountrypop <- scountrypop[-c(counter),]
  } else {
      counter <- counter+1
      break
  }
  }
}
scountrypop

#this creates a dataframe with only the movies that were produced in one of the relevant countries

minutes <- c()

for (i in 1:1555){
  addd <- unlist(str_split(scountrypop$duration[i], " "))
  minutes <- append(minutes, addd[1])
}


scountrypop$runtime <- minutes
scountrypop
#here i created a new column with the runtimes but in integers

showsc<- data.frame(scountrypop$country, scountrypop$runtime)

#this new dataframe has just the countries and their respective movie runtimes

showsc
```

This next chunk will contain the data visualization for this data


```{r}
mcouvrun <- ggplot(data = mcountrypop, aes(x = country, y = as.numeric(runtime), fill = country)) + geom_boxplot() + coord_flip() + theme(legend.position="none") + scale_fill_brewer(palette="BrBG")

mcouvrun + ggtitle("Movie Country of Production Boxplot") + xlab("Country") + ylab("Runtime")

#makes boxplot

moviesc
tapply(as.numeric(moviesc$mcountrypop.runtime), moviesc$mcountrypop.country, summary)
#applys the summary for each country


scouvrun <- ggplot(data = scountrypop, aes(x = country, y = as.numeric(runtime), fill = country)) + geom_boxplot() + coord_flip() + theme(legend.position="none") + scale_fill_brewer(palette="BrBG")

scouvrun + ggtitle("Movie Country of Production Boxplot") + xlab("Country") + ylab("Runtime")

#makes boxplot2

showsc
tapply(as.numeric(showsc$scountrypop.runtime), showsc$scountrypop.country, summary)
#applys the summary for each country2
```

In this next chunk I will now move onto the next variable: tv rating. Because there are not hundreds of ratings, i can skip the part where i find the frequency of each and use the most popular, instead i can work with all the ratings data

```{r}
mrating <- movies
#mrating

valid<- c("G", "NR", "PG", "PG-13", "R", "TV-14", "TV-G", "TV-MA", "TV-PG", "TV-Y","TV-Y7")

counter <- 1
for (i in 1:6131){
  if (!(movies$rating[i] %in%valid)){
    mrating <- mrating[-c(counter),]
  } else {
    counter <- counter+1
  }
}
mrating

#this gets rid of weird ratings

minutes <- c()

for (i in 1:6115){
  addd <- unlist(str_split(mrating$duration[i], " "))
  minutes <- append(minutes, addd[1])
}


mrating$runtime <- minutes
mrating
#here i created a new column with the runtimes but in integers

moviesr<- data.frame(mrating$rating, mrating$runtime)
#this new dataframe has just the rating and their respective movie runtimes
moviesr

srating <- shows
#srating

minutes <- c()

for (i in 1:2676){
  addd <- unlist(str_split(srating$duration[i], " "))
  minutes <- append(minutes, addd[1])
}


srating$runtime <- minutes
srating
#here i created a new column with the runtimes but in integers

showsr<- data.frame(srating$rating, srating$runtime)
#this new dataframe has just the rating and their respective movie runtimes

showsr

```
Now that I have the useable data, I can do visualization again

```{r}

rbar <- ggplot(mrating, aes(x=rating, fill = rating)) + geom_bar() + scale_fill_brewer(palette="BrBG")

rbar + ggtitle("Movies Per Rating") + xlab("TV Rating") + ylab("Number of Movies")

#makes barplot

rbar2 <- ggplot(srating, aes(x=rating, fill = rating)) + geom_bar() + scale_fill_brewer(palette="BrBG")

rbar2 + ggtitle("Shows Per Rating") + xlab("TV Rating") + ylab("Number of Shows")

#makes barplot2

mratvrun <- ggplot(data = moviesr, aes(x = mrating.rating, y = as.numeric(mrating.runtime), fill = mrating.rating)) + geom_boxplot() + theme(legend.position="none") + scale_fill_brewer(palette="BrBG")

mratvrun + ggtitle("Movie Rating Boxplot") + xlab("Rating") + ylab("Runtime")

#makes boxplot

moviesr
tapply(as.numeric(moviesr$mrating.runtime), moviesr$mrating.rating, summary)

#applys the summary for each rating2

sratvrun <- ggplot(data = showsr, aes(x = srating.rating, y = as.numeric(srating.runtime), fill = srating.rating)) + geom_boxplot() + theme(legend.position="none") + scale_fill_brewer(palette="BrBG")

sratvrun + ggtitle("Show Rating Boxplot") + xlab("Rating") + ylab("Runtime")

#makes boxplot2

showsr
tapply(as.numeric(showsr$srating.runtime), showsr$srating.rating, summary)

#applys the summary for each rating2
```
These next chunks will be used to analyze the next qualitative factor: the genre. Unlike the last factor, there are too many genres to consider all of them so once again I will first find those that are most popular and make datasets with only those. I will use the same approach I used for countries

```{r}
mgenre_list <-c()
for (i in 1:6131){
  addd <- unlist(str_split(movies$listed_in[i], ", "))
  mgenre_list <- append(mgenre_list, addd)
}
mgenre_list_df <- data.frame(mgenre_list)
#mgenre_list_df

mcountg <- count(mgenre_list_df, mgenre_list)
mcountg <- mcountg[order(-mcountg$n), ]
mcountg

#this creates a dataframe with each genre individually and then creates a second dataframe that counts their occurances

sgenre_list <-c()
for (i in 1:6131){
  addd <- unlist(str_split(shows$listed_in[i], ", "))
  sgenre_list <- append(sgenre_list, addd)
}
sgenre_list_df <- data.frame(sgenre_list)
#sgenre_list_df

scountg <- count(sgenre_list_df, sgenre_list)
scountg <- scountg[order(-scountg$n), ]
scountg

#this does the same thing but for the tv shows

mtopg <- c("International Movies", "Dramas", "Comedies", "Documentaries", "Action & Adventure", "Independent Movies", "Children & Family Movies", "Romantic Movies", "Thrillers", "Music & Musicals")

#list of genres with the most movies

mgenrepop <- movies
mtopg
counter <-1

for (i in 1:6131){
  for (j in movies$listed_in[i]){
    if (!(movies$listed_in[i] %in% mtopg)){
      mgenrepop <- mgenrepop[-c(counter),]
  } else {
      counter <- counter+1
      break
  }
  }
}
mgenrepop

#this creates a dataframe with only the movies that were listed as one of the relevant genres

minutes <- c()

for (i in 1:1032){
  addd <- unlist(str_split(mgenrepop$duration[i], " "))
  minutes <- append(minutes, addd[1])
}


mgenrepop$runtime <- minutes
mgenrepop
#here i created a new column with the runtimes but in integers

moviesg<- data.frame(mgenrepop$listed_in, mgenrepop$runtime)

#this new dataframe has just the countries and their respective movie runtimes

moviesg

```
this chunk will do the same as the last but for shows

```{r}
stopg <- c("International TV Shows", "TV Dramas", "TV Comedies", "Crime TV Shows", "Kids' TV", "Docuseries", "Romantic TV Shows", "Reality TV", "British TV Shows")

#list of genres with the most shows

sgenrepop <- shows
stopg
counter <-1

for (i in 1:2676){
  for (j in shows$listed_in[i]){
    if (!(shows$listed_in[i] %in% stopg)){
      sgenrepop <- sgenrepop[-c(counter),]
  } else {
      counter <- counter+1
      break
  }
  }
}
sgenrepop

#this creates a dataframe with only the movies that were listed as one of the relevant genres

minutes <- c()

for (i in 1:506){
  addd <- unlist(str_split(sgenrepop$duration[i], " "))
  minutes <- append(minutes, addd[1])
}


sgenrepop$runtime <- minutes
sgenrepop
#here i created a new column with the runtimes but in integers

showsg<- data.frame(sgenrepop$listed_in, sgenrepop$runtime)

#this new dataframe has just the genres and their respective movie runtimes

showsg
```

now i can do some data visulaization

```{r}
mgenvrun <- ggplot(data = mgenrepop, aes(x = mgenrepop.listed_in, y = as.numeric(runtime), fill = listed_in)) + geom_boxplot() + coord_flip() + theme(legend.position="none") + scale_fill_brewer(palette="BrBG")

mcouvrun + ggtitle("Movie Genre Boxplot") + xlab("Genre") + ylab("Runtime")

#makes boxplot

moviesc
tapply(as.numeric(moviesg$mgenrepop.runtime), moviesg$mgenrepop.listed_in, summary)
#applys the summary for each genre


sgenvrun <- ggplot(data = sgenrepop, aes(x = sgenrepoplisted_in, y = as.numeric(runtime), fill = listed_in)) + geom_boxplot() + coord_flip() + theme(legend.position="none") + scale_fill_brewer(palette="BrBG")

scouvrun + ggtitle("Show Genre Boxplot") + xlab("Genre") + ylab("Runtime")

#makes boxplot2

showsg
tapply(as.numeric(showsg$sgenrepop.runtime), showsg$sgenrepop.listed_in, summary)
#applys the summary for each country2
```

Moving onto quantitative varibles, I will begin by exploring to see if there exists some correlation between the release year and runtime. i can use all the data here

```{r}
mrelease <- movies
mrelease

minutes <- c()

for (i in 1:6131){
  addd <- unlist(str_split(mrelease$duration[i], " "))
  minutes <- append(minutes, addd[1])
}


mrelease$runtime <- minutes
mrelease
#here i created a new column with the runtimes but in integers

moviesre<- data.frame(mrelease$release_year, mrelease$runtime)
#this new dataframe has just the release yr and their respective movie runtimes
moviesre

srelease <- shows
#srelease

minutes <- c()

for (i in 1:2676){
  addd <- unlist(str_split(srelease$duration[i], " "))
  minutes <- append(minutes, addd[1])
}


srelease$runtime <- minutes
srelease
#here i created a new column with the runtimes but in integers

showsre<- data.frame(srelease$release_year, srelease$runtime)
#this new dataframe has just the release yr and their respective movie runtimes

showsre
```

Now i can use the data to make a visualization of it. For the quanitative data i will primarily use scatterplots.

```{r}
mrelvrun <- ggplot(data = moviesre, aes(x = mrelease.release_year, y = as.numeric(mrelease.runtime))) + geom_point()

mrelvrun + ggtitle("Movie Release Year Scatterplot") + xlab("Year") + ylab("Runtime")

#makes scatterplot1

srelvrun <- ggplot(data = showsre, aes(x = srelease.release_year, y = as.numeric(srelease.runtime))) + geom_point()

srelvrun + ggtitle("Show Release Year Scatterplot") + xlab("Year") + ylab("Runtime")

#makes scatterplot2

recount <- count(movies, release_year)
recount <- recount[order(-recount$n), ]
recount

#creates a dataframe with the number of movies released per year

mrelvct <- ggplot(recount, aes(x = release_year, y = n)) + geom_line()

mrelvct + ggtitle("Movies per Year") + xlab("Year") + ylab("Number of Movies")

#creates lineplot1

recount2 <- count(shows, release_year)
recount2 <- recount[order(-recount$n), ]
recount2

#creates a dataframe with the number of movies released per year

mrelvct <- ggplot(recount2, aes(x = release_year, y = n)) + geom_line()

mrelvct + ggtitle("Shows per Year") + xlab("Year") + ylab("Number of Movies")

#creates lineplot2
```

Moving onto the next quntitative factor. For date added to netflix, because the data comes in the form of "month day, year", i first have to convert the data into a numeric form by making each date to follow the form "year.decimal" by considering the proprtion of the year each date has passed, I will make a function to do this in this chunk

```{r}
movies

madded <- c()
for (i in 1:6131){
  madded <- append(madded, movies$date_added[i])
}

#makes a vector containing the dates added

date_to_prop <- function(cha){
  use <-c()
  dec <- c()
  prop <- (0)
  use <- unlist(strsplit(cha, ", "))
  dec <- unlist(strsplit(use[1], " ")) 
  
  if ((dec[1] == "January") | (is.na(dec[1]))){
    prop <- prop
  }else if(dec[1] == "February"){
    prop <- prop+1/12
  }else if(dec[1] == "March"){
    prop <- prop+2/12
  }else if(dec[1] == "April"){
    prop <- prop+3/12
  }else if(dec[1] == "May"){
    prop <- prop+4/12
  }else if(dec[1] == "June"){
    prop <- prop+5/12
  }else if(dec[1] == "July"){
    prop <- prop+6/12
  }else if(dec[1] == "August"){
    prop <- prop+7/12
  }else if(dec[1] == "September"){
    prop <- prop+8/12
  }else if(dec[1] == "October"){
    prop <- prop+9/12
  }else if(dec[1] == "November"){
    prop <- prop+10/12
  }else if(dec[1] == "December"){
    prop <- prop+11/12
  }
  
  day<- as.numeric(dec[2])
  year<- as.numeric(use[2])
  prop <- prop + day/365 + year
  prop
}
#date_to_prop("November 3, 2021")

#creates function that will turn character date into numeric data

madded_prop <- c()
for (i in 1:6131){
  addd <- date_to_prop(madded[i])
  madded_prop <- append(madded_prop, addd)
}

#creates new vector with numeric date data

sadded <- c()
for (i in 1:2676){
  sadded <- append(sadded, shows$date_added[i])
}

#makes a vector containing the dates added, this time for shows

sadded_prop <- c()
for (i in 1:2676){
  addd <- date_to_prop(sadded[i])
  sadded_prop <- append(sadded_prop, addd)
}

#creates new vector with numeric date data for shows

madded_df <- data.frame(madded_prop, mrelease$runtime)
madded_df

sadded_df <- data.frame(sadded_prop, srelease$runtime)
sadded_df

#makes movie and show df
```

now that i have converted the data into something quantifiably usable andorganized it into dataframes, i can move onto the data visualization

```{r}
maddvrun <- ggplot(data = madded_df, aes(x = madded_prop, y = as.numeric(mrelease.runtime))) + geom_point()

maddvrun + ggtitle("Movie Added Date Scatterplot") + xlab("Date (by Year)") + ylab("Runtime")

#makes scatterplot1

saddvrun <- ggplot(data = sadded_df, aes(x = sadded_prop, y = as.numeric(srelease.runtime))) + geom_point()

saddvrun + ggtitle("Show Added Date Scatterplot") + xlab("Date (by Year)") + ylab("Runtime")

#makes scatterplot2

```
onto the next quantitative factor: time of year added to netflix. this will use the same appraoch as the last as i turn each date into numeric data. I will have to revamp my function to not include the year this time
```{r}

movies

maddedy <- c()
for (i in 1:6131){
  maddedy <- append(maddedy, movies$date_added[i])
}

#makes a vector containing the dates added

date_to_propy <- function(cha){
  use <-c()
  dec <- c()
  prop <- (0)
  use <- unlist(strsplit(cha, ", "))
  dec <- unlist(strsplit(use[1], " ")) 
  
  if ((dec[1] == "January") | (is.na(dec[1]))){
    prop <- prop
  }else if(dec[1] == "February"){
    prop <- prop+1/12
  }else if(dec[1] == "March"){
    prop <- prop+2/12
  }else if(dec[1] == "April"){
    prop <- prop+3/12
  }else if(dec[1] == "May"){
    prop <- prop+4/12
  }else if(dec[1] == "June"){
    prop <- prop+5/12
  }else if(dec[1] == "July"){
    prop <- prop+6/12
  }else if(dec[1] == "August"){
    prop <- prop+7/12
  }else if(dec[1] == "September"){
    prop <- prop+8/12
  }else if(dec[1] == "October"){
    prop <- prop+9/12
  }else if(dec[1] == "November"){
    prop <- prop+10/12
  }else if(dec[1] == "December"){
    prop <- prop+11/12
  }
  
  day<- as.numeric(dec[2])
  year<- as.numeric(use[2])
  prop <- prop + day/365
  prop
}
date_to_propy("November 3, 2021")

#creates function that will turn character date into numeric data

madded_propy <- c()
for (i in 1:6131){
  addd <- date_to_propy(maddedy[i])
  madded_propy <- append(madded_propy, addd)
}

#creates new vector with numeric date data

saddedy <- c()
for (i in 1:2676){
  saddedy <- append(saddedy, shows$date_added[i])
}

#makes a vector containing the dates added, this time for shows

sadded_propy <- c()
for (i in 1:2676){
  addd <- date_to_propy(saddedy[i])
  sadded_propy <- append(sadded_propy, addd)
}

#creates new vector with numeric date data for shows

maddedy_df <- data.frame(madded_propy, mrelease$runtime)
maddedy_df

saddedy_df <- data.frame(sadded_propy, srelease$runtime)
saddedy_df

#makes movie and show df

maddedy2 <- c()
for (i in maddedy){
  use <-c()
  dec <- c()
  use <- unlist(strsplit(i, ", "))
  dec <- unlist(strsplit(use[1], " "))
  maddedy2 <- append(maddedy2, dec[1])
}

mmonth_df <- data.frame(maddedy2)
mmonth_df

saddedy2 <- c()
for (i in saddedy){
  use <-c()
  dec <- c()
  use <- unlist(strsplit(i, ", "))
  dec <- unlist(strsplit(use[1], " "))
  saddedy2 <- append(saddedy2, dec[1])
}

smonth_df <- data.frame(saddedy2)
smonth_df

#creates dataframe to count the frequencies of each month
```

i will now make scatterplots for this data as well. i will also generalize by month to make histograms of listings per month

```{r}
madyvrun <- ggplot(data = maddedy_df, aes(x = madded_propy, y = as.numeric(mrelease.runtime))) + geom_point()

madyvrun + ggtitle("Movie Added Date Scatterplot") + xlab("Date (by Month)") + ylab("Runtime")

#makes scatterplot1

sadyvrun <- ggplot(data = saddedy_df, aes(x = sadded_propy, y = as.numeric(srelease.runtime))) + geom_point()

sadyvrun + ggtitle("Show Added Date Scatterplot") + xlab("Date (by Month)") + ylab("Runtime")

#makes scatterplot2

macount <- count(mmonth_df, maddedy2)
macount <- macount[order(-macount$n), ]
macount

sacount <- count(smonth_df, saddedy2)
sacount <- sacount[order(-sacount$n), ]
sacount

#makes dataframes with the individual counts of listings per month

ggplot(mmonth_df, aes(x = maddedy2)) + geom_histogram(stat = "count") + ggtitle("Movie Added Date Histogram") + xlab("Date (by Month)") + ylab("Count")


ggplot(smonth_df, aes(x = saddedy2)) + geom_histogram(stat = "count") + ggtitle("Show Added Date Histogram") + xlab("Date (by Month)") + ylab("Count")

#makes histograms
```

